---
title: "dissertation"
output:
  html_document: default
  pdf_document: default
---

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
#install.packages("jtools")  
#install.packages("interactions")
#install.packages("emmeans")
library(dplyr)
library(tidyr)
library(purrr)
library(lme4)
library(tibble)
library(lmerTest)
library(data.table)
library(ggplot2)
library(interactions)
library(jtools)           
library(mvtnorm)
library(emmeans)
```

```{r setup, include=FALSE}
df <- read.csv('/Users/wenyigu/Desktop/2023-07-21_DummyCodedMerged_LIWC_withProductivityData.csv')

df<-df%>%
  mutate(
    promotion_to_associate_professor_candidates = ifelse(action_s_requested == "Promotion to Associate Professor", 1,0),
    promotion_to_full_professor_candidates = ifelse(action_s_requested == "Promotion to Professor",1,0),
    ten.ext.nc = ifelse(tenure_extension_requested_non_covid_related_coded_as_1_no_and_2_yes == 2, 1, 0),
    ten.ext.c = ifelse(covid_related_tenure_extension_coded_as_1_no_and_2_yes == 2, 1, 0),
    mean.can.hi = mean(h_index_at_pt_candidate, na.rm = T),
    mean.can.tc = mean(total_citations_at_pt_candiate, na.rm = T),
    can.fem = ifelse(candidate_gender_num == 2, 1, 0),
    lw.fem = ifelse(lw_gender_num == 2, 1, 0),
    lw.urm = ifelse(letter_writer_urm_num == 2, 1, 0),
    can.asian = ifelse(candidate_ethnicity_num == 2, 1, 0),
    can.white = ifelse(candidate_ethnicity_num == 1, 1, 0),
    can.hisp = ifelse(candidate_ethnicity_num == 3, 1, 0),
     can.black = ifelse(candidate_ethnicity_num == 4, 1, 0),
    can.asianvswhite = as.numeric(case_when(candidate_ethnicity_num == 1 ~ 0, candidate_ethnicity_num == 2 ~ 1)),
    can.hispvswhite = as.numeric(case_when(candidate_ethnicity_num == 1 ~ 0, candidate_ethnicity_num == 3 ~ 1)),
    can.blackvswhite = as.numeric(case_when(candidate_ethnicity_num == 1 ~ 0, candidate_ethnicity_num == 4 ~ 1)),
    tenure_approved = case_when(president_vote == 1 ~ 0, 
                                   president_vote == 2 ~ 1,
                                   is.na(president_vote) == T & provost_vote_1_no_2_yes == 1 ~ 0,
                                   is.na(president_vote) == T & provost_vote_1_no_2_yes == 2 ~ 1,
                                   is.na(president_vote) == T & is.na(provost_vote_1_no_2_yes) == T & dean_vote_1_no_2_yes == 1 ~ 0,
                                    is.na(president_vote) == T & is.na(provost_vote_1_no_2_yes) == T & dean_vote_1_no_2_yes == 2 ~ 1,
                                    is.na(president_vote) == T & is.na(provost_vote_1_no_2_yes) == T & is.na(dean_vote_1_no_2_yes) == T & department_chair_vote_1_no_2_yes == 1 ~ 0,
                                    is.na(president_vote) == T & is.na(provost_vote_1_no_2_yes) == T & is.na(dean_vote_1_no_2_yes) == T & department_chair_vote_1_no_2_yes == 2 ~ 1),
    can.nonURM = ifelse(candidate_ethnicity_num == 2 | candidate_ethnicity_num == 1, 1, 0),
    can.URM = ifelse(candidate_ethnicity_num == 3 | candidate_ethnicity_num == 4, 1, 0))


#df1 <- df %>%
 # filter(school_code %in% c(3,4,9))

df<- df%>%
  group_by(candidate_code)%>%
  mutate(
    can.hi.center = h_index_at_pt_candidate - mean.can.hi,
    can.tc.center = total_citations_at_pt_candiate - mean.can.tc,
    candidate_discipline_cip_code = as.factor(candidate_discipline_cip_code))

df <- df %>%
  mutate(UH = ifelse(school_code ==1, 1, 0),
         Hampton = ifelse(school_code == 2,1, 0),
         TAMU = ifelse(school_code == 3, 1, 0), #
         LSU = ifelse(school_code == 4, 1, 0), #
         Rice = ifelse(school_code == 5, 1, 0),
         Lehigh = ifelse(school_code == 6, 1, 0),
         WPI = ifelse(school_code == 7, 1, 0),
         UTRGV = ifelse(school_code == 8, 1, 0),
         UA = ifelse(school_code == 9, 1, 0), #
         TSU = ifelse(school_code == 10, 1, 0))
                  
promotion_to_associate_professor_candidates <- df %>% filter(action_s_requested %in% c("Promotion to Associate Professor", "Promotion to Associate Professor without tenure"))
promotion_to_full_professor_candidates <- df %>% filter(action_s_requested == "Promotion to Professor")

#sum ERL per school
count_lw_school <- df %>%
  group_by(school_code) %>%
  summarize(count_lw_code = sum(!is.na(lw_code)))
count_lw_school

#sum candidates per school
total_candidates <- df %>%
    group_by(school_code) %>%
  summarise(count_candidates = n_distinct(candidate_code))
total_candidates

#sum candidates per race
total_candidates_race <- df %>%
    group_by(candidate_ethnicity_num) %>%
  summarise(count_candidates = n_distinct(candidate_code))
total_candidates_race

# % of dei language by race
count_DEI_race <- df %>%
  group_by(candidate_ethnicity_num) %>%
  summarize(count_DEI_race = mean(dei,na.rm = T))
count_DEI_race

# % of dei language by URM status
count_DEI_URM <- df %>%
  group_by(can.URM) %>%
  summarize(count_DEI_URM = mean(dei,na.rm = T))

##summary stats
summary_statistics <- summary(df[c("clout", "analytic", "authentic", "tone", "dei", "h_index_of_candidate", "can.URM", "department_negative_vote_percentage", "college_committee_negative_vote_percentage","tenure_approved")])

##mean values
mean_values <- sapply(df[c("clout", "analytic", "authentic", "tone", "dei", "h_index_of_candidate", "can.URM", "department_negative_vote_percentage", "college_committee_negative_vote_percentage","tenure_approved")], mean, na.rm = TRUE)

##sd values
sd_values <- sapply(df[c("clout", "analytic", "authentic", "tone", "dei", "h_index_of_candidate", "can.URM", "department_negative_vote_percentage", "college_committee_negative_vote_percentage","tenure_approved")], sd, na.rm = TRUE)

##CORRELATION TABLE WITH PVALUE
# Assuming df is your data frame
variables_of_interest <- df[c("tone", "analytic", "clout", "authentic", "dei", "h_index_of_candidate", "can.URM", "department_negative_vote_percentage", "college_committee_negative_vote_percentage","tenure_approved")]

# Convert variables to numeric
numeric_variables <- apply(variables_of_interest, 2, as.numeric)

# Calculate the correlation matrix
correlation_matrix <- cor(numeric_variables, use = "complete.obs")

# Function to calculate correlation coefficients and p-values
cor_test <- function(x, y) {
  test_result <- cor.test(x, y, use = "complete.obs")
  return(test_result$p.value)
}

# Apply the function to all pairs of variables in the correlation matrix
p_values_matrix <- outer(
  colnames(correlation_matrix),
  colnames(correlation_matrix),
  Vectorize(function(x, y) cor_test(numeric_variables[, x], numeric_variables[, y]))
)
# Display the correlation matrix
correlation_matrix
# Display the matrix of p-values
p_values_matrix
```

### URM ANALYSES 
### MLM 
```{r}
#URM more likely to have DEI language
df$candidate_discipline_cip_code <- as.numeric(as.character(df$candidate_discipline_cip_code))

URM.dei.results<- lmer(dei ~ can.hi.center + can.URM + candidate_discipline_cip_code + school_code +promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates + ( 1 | candidate_code), data = df)
summary(URM.dei.results)


#nonURM less likely to have DEI language
nonURM.dei.results<- lmer(dei ~ can.hi.center + can.nonURM + candidate_discipline_cip_code + school_code +promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates + ( 1 | candidate_code), data = df)
summary(nonURM.dei.results)
```

# Do Linguistic Features Predict Voting Outcomes (URM status as moderator)
# Negative voting %
```{r}
# Convert "can.nonURM" to a factor variable (if it's not already a factor)
df$can.nonURM <- factor(df$can.nonURM)

# Set up contrast coding to compare each group to "can.nonURM" (reference group)
contrasts(df$can.nonURM) <- contr.treatment(levels(df$can.nonURM))

##DEPT not sig
urm.dei.dnvp.results <- glm(department_negative_vote_percentage ~ dei + (can.URM) +  dei * (can.URM) + h_index_at_pt_candidate + school_code + candidate_discipline_cip_code + promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates, data = df)
print(summary(urm.dei.dnvp.results))

##COLLEGE (+)
urm.dei.cnvp.results <- glm(college_committee_negative_vote_percentage ~ dei + (can.URM)+ dei*(can.URM) + h_index_at_pt_candidate + school_code+ candidate_discipline_cip_code+ promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates, data = df)
print(summary(urm.dei.cnvp.results))
interact_plot(urm.dei.cnvp.results, pred = dei, modx = can.URM)

##CONTROL ##COLLEGE (+)
urm.dei.cnvp.results <- glm(college_committee_negative_vote_percentage ~ dei + (can.URM)+ dei*(can.URM) + h_index_at_pt_candidate + school_code+ candidate_discipline_cip_code+ promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates + analytic + clout + tone + authentic , data = df)
print(summary(urm.dei.cnvp.results))
interact_plot(urm.dei.cnvp.results, pred = dei, modx = can.URM)

#emm and simple slope for URM
m_dei<- mean(df$dei, na.rm = TRUE)
sd_dei<- sd(df$dei, na.rm = TRUE)
emm <- emmeans(urm.dei.cnvp.results,  ~ dei*can.URM,
  cov.keep = 3, at = list(
  dei = c(m_dei-sd_dei, m_dei, m_dei+sd_dei)), level = 0.95)
summary(emm)
simpleSlope <- emtrends(urm.dei.cnvp.results, pairwise ~ can.URM, var = "dei", level = 0.95)
summary(simpleSlope)

##UNI (+)
#urm.dei.unvp.results <- glm(university_committee_negative_vote_percentage ~ dei + (can.URM)+ dei*(can.URM) + h_index_at_pt_candidate + school_code+ candidate_discipline_cip_code+ promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates, data = df)
#print(summary(urm.dei.unvp.results))
#interact_plot(urm.dei.unvp.results, pred = dei, modx = can.URM)

#emm and simple slope for URM
#emm <- emmeans(urm.dei.unvp.results,  ~ dei*can.URM,
 # cov.keep = 3, at = list(
  #dei = c(m_dei-sd_dei, m_dei, m_dei+sd_dei)), level = 0.95)
#summary(emm)
#simpleSlope <- emtrends(urm.dei.unvp.results, pairwise ~ can.URM, var = "dei", level = 0.95)
#summary(simpleSlope)
```

# Do Linguistic Features Predict Voting Outcomes (URM status as moderator)
# unanimous votes
```{r}
##DEPT not sig
urm.dei.duv.results <- glm(department_unanimous_vote_abstentions_included ~ dei + (can.URM)+ dei*(can.URM) + h_index_at_pt_candidate + school_code+ candidate_discipline_cip_code+ promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates, data = df)
print(summary(urm.dei.duv.results))

##COLLEGE not sig
urm.dei.cuv.results <- glm(college_committee_unanimous_vote_abstentions_included ~ dei + (can.URM)+ dei*(can.URM) + h_index_at_pt_candidate + school_code+ candidate_discipline_cip_code+ promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates, data = df)
print(summary(urm.dei.cuv.results))

## UNI not sig
urm.dei.uuv.results <- glm(university_committee_unanimous_vote_abstentions_included ~ dei + (can.URM)+ dei*(can.URM) + h_index_at_pt_candidate + school_code+ candidate_discipline_cip_code+ promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates, data = df)
print(summary(urm.dei.uuv.results))
```

# Do Linguistic Features Predict Voting Outcomes (URM status as moderator)
# positive votes
```{r}
##DEPT not sig
urm.dei.dpv.results <- glm(department_positive_vote ~ dei + (can.URM)+ dei*(can.URM) + h_index_at_pt_candidate + school_code+ candidate_discipline_cip_code + promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates, data = df)
print(summary(urm.dei.dpv.results))

##COLLEGE not sig
urm.dei.cpv.results <- glm(college_committee_positive_vote ~ dei + (can.URM)+ dei*(can.URM) + h_index_at_pt_candidate + school_code+ candidate_discipline_cip_code+ promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates, data = df)
print(summary(urm.dei.cpv.results))

##UNI not sig
urm.dei.upv.results <- glm(university_committee_positive_vote ~ dei + (can.URM)+ dei*(can.URM) + h_index_at_pt_candidate + school_code+ candidate_discipline_cip_code+ promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates, data = df)
print(summary(urm.dei.upv.results))
```

# Do Linguistic Features Predict Voting Outcomes (URM status as moderator)
# Provost vote
```{r}
## (-)
urm.dei.ta.results <- glm(tenure_approved ~ dei + (can.URM)+ dei*(can.URM) + h_index_at_pt_candidate + school_code+ candidate_discipline_cip_code+ promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates, data = df)
print(summary(urm.dei.ta.results))
interact_plot(urm.dei.ta.results, pred = dei, modx = can.URM)

## CONTROL
urm.dei.ta.results <- glm(tenure_approved ~ dei + (can.URM)+ dei*(can.URM) + h_index_at_pt_candidate + school_code+ candidate_discipline_cip_code+ promotion_to_associate_professor_candidates + promotion_to_full_professor_candidates + authentic + analytic + clout + tone, data = df)
print(summary(urm.dei.ta.results))
interact_plot(urm.dei.ta.results, pred = dei, modx = can.URM)

#emm and simple slope for URM
emm <- emmeans(urm.dei.ta.results,  ~ dei*can.URM,
  cov.keep = 3, at = list(
  dei = c(m_dei-sd_dei, m_dei, m_dei+sd_dei)), level = 0.95)
summary(emm)
simpleSlope <- emtrends(urm.dei.ta.results, pairwise ~ can.URM, var = "dei", level = 0.95)
summary(simpleSlope)

```